---
title: "Linear Models ATS Homework"
author: "ATS"
date: "2025-04-2"
output:
  md_document:
    variant: markdown_github
  html_document: default
---

## Packages needed for these notes:

```{r}
library(tidyverse)
library(lme4)
# install.packages("packages/emmeans_1.10.6.tar.gz", repos = NULL, type = "source")
library(emmeans)
# install.packages("packages/multcomp_1.4-26.tar.gz", repos = NULL, type = "source")
library(multcomp)
# install.packages("multcompView")
library(multcompView)
```



# Introduction to regression analysis in R


y(hat) = β0 + β1x

where y(hat) = equation 
  - β0 = intercept
  - β1 = slope 
  - x = x points

The actual modeled line is equal to the equation adds on the error term.
yi = β0 + β1xi + ε~N(0,σ)


where yi (y sub i) = each point on the y axis 
  - β0 = intercept
  - β1 = slope 
  - xi = each x point 
  - then ε = error
  - and N(0,σ) = normal distribution with:
    - mean = 0 
    - standard deviation = sigma

The goal of all of this is to estimate three terms. 
  - slope
  - intercept
  - standard deviation4
 
 *The smaller the SSE and higher the SSR, the smaller the p-value*

*Then, we want to know if it is statistically significant. What is the chance we get the result? - that is your p-value.*

 **Connect in your brain which test I will need to use for my data**



## cont<sup>s</sup> X & cont<sup>s</sup> Y

Loading in cars database as example

```{r}
data("mtcars")
```

The variables inside the cars dataset include: 

  - mpg = the mpg
  - cyl = number of cylinders
  - disp = displacement
  - hp = gross horsepower
  - drat = rear axle ratio
  - wt = weight
  - qsec = quarter mile time 
  - vs = engine shape V = 0 or Straight = 1
  - am = Transmission type; 0 = auto, 1 = manual
  - gear = number of forward gears
  - carb = number of carburetors

If you have two continuous variables you will want a scatter plot

#### Weight vs MPG
```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = lm, se = FALSE, color = "grey") +
  geom_point(aes(color = wt)) +
  xlab("Weight") + 
  ylab("Miles per gallon") +
  scale_colour_gradient(low = "forestgreen", high = "black") +
  theme_classic()
```


This looks to be a general negative relationship, but how can we be sure?

### linear models

```{r}
lm(mpg~wt, data = mtcars)
```


```{r}
lm1 <- lm(mpg~wt, data = mtcars)
summary(lm1)
```


## Anovas

Analysis of variance is used for catagorical variables, but can be used on linear models as well.



```{r}
anova(lm1)
```

## Correlation Test

```{r}
cor.test(mtcars$mpg,mtcars$wt)
```


## Assumptions

In general, there are several assumptions in a regression, linear model, ANOVA, or whatever you want to call it.

They are:

- y is continuous
- error is normally distributed 
- relationship is linear 
- homoskedasticity 
- sigma is consistent 
- independent samples


### Getting residuals from model above

```{r}
model <- lm(mpg~wt, data = mtcars)

ggplot(model, aes(y = .resid, x = .fitted)) +
  geom_point() +
  geom_hline(yintercept = 0)
```



## New Dataset - Bull Richness


```{r}
bull.rich <- read.csv("../data/Bull_richness.csv")
```


Demonstrating: 
- Anova w/ two groups (ttest)

- Anova w/ more than one category MANOVA



```{r}
bull.rich.sub <- bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.") 
 
  
t.test(richness ~ Fungicide, data = bull.rich.sub)

```

```{r}
summary(lm(richness ~ Fungicide, data = bull.rich.sub))

```


## Multiple Categories





subset data


```{r}
bull.rich.sub2 <- bull.rich %>%
  filter(Fungicide == "C" & Treatment == "Conv." & Crop == "Corn")
```



```{r}
ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot()
```

let's move the bars so the factors are in ascending order

```{r}
bull.rich.sub2$GrowthStage <- factor(bull.rich.sub2$GrowthStage, levels = c("V6", "V8", "V15"))

ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot()
```


```{r}
lm3 <- lm(richness ~ GrowthStage, data = bull.rich.sub2)
summary(lm3)
```


same p-value w


```{r}
anova(lm3)
```



## emmeans and multcomp packages


```{r}
lsmeans <- emmeans(lm3, ~GrowthStage) # estimates least-square means of variety within site vs year



Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey adjustment by default. 
Results_lsmeans
  
```

## Interaction Terms:
The interactions between factors. In other words: what if the effect of one factor depends on the other?


```{r}
bull.rich.sub3 <- bull.rich %>%
  filter(Treatment == "Conv." & Crop == "Corn")

bull.rich.sub3$GrowthStage <- factor(bull.rich.sub3$GrowthStage, levels = c("V6", "V8", "V15"))
```






```{r}
# write it like this
lm.inter <- lm(richness ~ GrowthStage + Fungicide + GrowthStage:Fungicide, data = bull.rich.sub3)

# or like this
lm.interB <-lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)


summary(lm.inter)
```


```{r}
anova(lm.inter)
```
Effect of fungicide depends on the growth stage





three seperate tests: one within each growth stage of the effect of fungicide.
Did not effect richness except v8

```{r}
lsmeans <- emmeans(lm.inter, ~Fungicide|GrowthStage) # estimate lsmeans of variety within siteXyear
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeans
```



```{r}
bull.rich.sub3 %>%
  ggplot(aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  geom_boxplot()
```


## Mixed Effects Models
All mixed-effects mdoels have a common fixed and random effects. 

*Common fixed effects*
- Treatment *
- Species *
- Genes

*Common random effects (blocking factor)*
- Year *
- Replicate * 
- Trial
- Individuals  *
- Fields *

(*) = possible effects for my data

```{r}
lm.inter2<-lmer(richness ~ GrowthStage*Fungicide + (1|Rep), data = bull.rich.sub3)
summary(lm.inter2)

```
(compare with above)
```{r}
summary(lm.inter)
```





