---
title: "Hw Notes Loops and Functions"
author: "ATS"
date: "2025-04-01"
output:
  md_document:
    variant: markdown_github
  html_document: default
---

Dependencies needed: 

```{r}

library(readr)
library(ggplot2)
library(drc) 
library(tidyverse)
library(dplyr)
library(drc)


```


## Anatomy of a function:

sample.function <- function(... variable goes here ...){
  .... code goes here.... 
  return(... output ...)
}


## Making a function

```{r}
F_to_C <- function(fahrenheit_temp){
  celsius <- (5*(fahrenheit_temp - 32)/9)
  return(celsius)
}

F_to_C(32)

```

### My attempt
using : F = C x (9/5)+32

```{r}
C_to_F <- function(cels_temp){
  fahrenheit <- (9/5)*(cels_temp)+32
  return(fahrenheit)
}

C_to_F(32)

```
### It worked!




# Iterations:


```{r}
rep("A", 3)
rep(c("A","B"),3)
rep(c("A","B","C"),each=3)
rep(c("A","B","C"),times=3)
rep(c("A","B","C"),each=3,times=3)


```


```{r}

seq(from = 1, to = 7)
seq(from = 0, to = 10, by = 2) 

rep(seq(from = 0, to = 10, by = 2), times = 3, each = 2)
LETTERS
seq_along(LETTERS[1:5])

```



#### The **Dreaded** for loop



```{r}
for (i in -30:100){
  result <- F_to_C(i)
  print(result)
}
```


## What if we wanted to do something with the result of the iteration?

1. Set a R object to NULL

2. Set your for loop

3. Save the result of your for loop into a dataframe each iteration

4. append one row of the dataframe to the null object each iteration of the loop. 



```{r}
celcius.df <- NULL 
for (i in -30:100){   
result_i <- data.frame(F_to_C(i), i)   
celcius.df <- rbind.data.frame(celcius.df, result_i) 
}

head(celcius.df)
```


#### More of a challenge

```{r}
EC50.data <- read.csv("../data/EC50_all.csv")

isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~ 
        EC50.data$conc[EC50.data$is == "ILSO_5-41c"], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]


```




```{r}
nm <- unique(EC50.data$is)

for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    EC50
}
```



```{r}
EC50.ll4 <- NULL # create a null object 
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    EC50
    isolate.ec_i <- data.frame(nm[[i]], EC50) # create a one row dataframe containing just the isolate name and the EC50
    colnames(isolate.ec_i) <- c("Isolate", "EC50") # change the column names
    
    # Then we need to append our one row dataframe to our null dataframe we created before
    # and save it as EC50.ll4. 
    EC50.ll4 <- rbind.data.frame(EC50.ll4, isolate.ec_i)
}

ggplot(EC50.ll4, aes(x = EC50)) + geom_histogram() + theme_classic() 
```



```{r}
EC50.data %>%
  group_by(is) %>%
  nest() %>%
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc, 
                              fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                         names = c("Slope", "Lower", "Upper", "EC50"))))) %>%
  mutate(ec50 = map(ll.4.mod, ~ED(., 
                              respLev = c(50), 
                              type = "relative",
                              interval = "delta")[[1]])) %>%
  unnest(ec50)
```


