---
title: "Linear Models ATS Homework"
author: "Alex Swystun & Sol Llerena"
date: "2025-04-2"
output:
  pdf_document: default
  md_document:
    variant: markdown_github
  html_document: default
  
---


1. 4 pts. Read in the data called `“PlantEmergence.csv”` using a relative file path and load the
following libraries. `tidyverse`, `lme4`, `emmeans`, `multcomp`, and `multcompView`. Turn the `Treatment` , `DaysAfterPlanting` and `Rep` into factors using the function as.factor

```{r}
library(readr)
Plant_Emergence <- read.csv("../data/PlantEmergence.csv")
library(tidyverse)
library(lme4)
#install.packages("estimability")
library(estimability)
#install.packages("packages/emmeans_1.10.6.tar.gz", repos = NULL, type = "source")
library(emmeans)
#install.packages("packages/multcomp_1.4-26.tar.gz", repos = NULL, type = "source")
library(multcomp)
#install.packages("multcompView")
library(multcompView)
library(ggplot2)


# Turn the Treatment , DaysAfterPlanting and Rep into factors using the function as.factor

Plant_Emergence$DaysAfterPlanting <- as.factor(Plant_Emergence$DaysAfterPlanting)
Plant_Emergence$Treatment <- as.factor(Plant_Emergence$Treatment)
Plant_Emergence$Rep <- as.factor(Plant_Emergence$Rep)


str(Plant_Emergence)
```


2. 5 pts. Fit a linear model to predict `Emergence` using `Treatment` and
`DaysAfterPlanting` along with the interaction. Provide the summary of the linear
model and ANOVA results.


```{r}

# We can make the model like this
lm1 <- lm(Emergence~Treatment + DaysAfterPlanting + Treatment:DaysAfterPlanting, data = Plant_Emergence)

# Or like this, these are the same thing:
lm1 <- lm(Emergence~Treatment * DaysAfterPlanting, data = Plant_Emergence)

summary(lm1)
anova(lm1)

```


3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

  Based on the results, no, it does not look like we need to fit the interaction term, as the interaction between `Treatment` and `DaysAfterPlanting` was not significant for the anova     .  



```{r}

# Or like this, these are the same thing:
lm2 <- lm(Emergence~Treatment + DaysAfterPlanting, data = Plant_Emergence)

summary(lm2)
anova(lm2)

```


The coefficient estimate for Treatment 2 was -134.531, therefore, the intercept for Treatment 2 would be *182.163 - 134.531 = 47.632* 

The slope of Treatment 2 in our model was negative with significance, which means the mean of plant emergence is negatively effected by Treatment 2 by an average of 47 units.


4. 5 pts. Calculate the least square means for `Treatment` using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.



```{r}


lsmeans <- emmeans(lm2, ~Treatment) # estimates least-square means of variety within site vs year

Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey adjustment by default. 
Results_lsmeans

```


 
Based on the Tukey comparisons, there are groups of Treatments that seem to be affecting the plant emergence in different ways. Groups 5, 3, and 6; 8, 4, 9, and 1; as well as individual groups 7 and 2 all have varying effects on plant emergence, but each group has similar effects.  





5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 



```{r}
plot_cldbars_onefactor <- function(lm2,Treatment) {
  data <- lm2$model
  variables <- colnames(lm2$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm2, as.formula(paste("~", "Treatment"))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.


   # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c("Treatment", "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm2$model %>%
    group_by(!!sym("Treatment")) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = "Treatment") %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym("Treatment"), y = !! sym("Emergence"))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character("Treatment")) +
    theme_classic()
  
  return(plot)
}
plot_cldbars_onefactor(lm2 = lm2, Treatment =  "Treatment")

```

Each bar has a letter representation for which group they are assigned, based on the effect that treatment has on the number of emerging plants. For example, only treatment two has the c group, as it is the only treatment with so few emerging plants, while other treatments share a letter "a" or "b" with another treatment.


6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.


[link to github](https://github.com/Aswystun/PLPA)
[link to this assignment](https://github.com/Aswystun/PLPA/tree/master/coding_challenge_7)



