---
title: "Linear Models ATS Homework"
author: "Alex Swystun & Sol Llerena"
date: "2025-04-2"
output:
  md_document:
    variant: markdown_github
  html_document: default
---


1. 4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the
following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting and Rep into factors using the function as.factor

```{r}
library(readr)
Plant_Emergence <- read.csv("../data/PlantEmergence.csv")
library(tidyverse)
library(lme4)
install.packages("estimability")
library(estimability)
install.packages("packages/emmeans_1.10.6.tar.gz", repos = NULL, type = "source")
library(emmeans)
install.packages("packages/multcomp_1.4-26.tar.gz", repos = NULL, type = "source")
library(multcomp)
install.packages("multcompView")
library(multcompView)


# Turn the Treatment , DaysAfterPlanting and Rep into factors using the function as.factor

Plant_Emergence$DaysAfterPlanting <- as.factor(Plant_Emergence$DaysAfterPlanting)
Plant_Emergence$Treatment <- as.factor(Plant_Emergence$Treatment)
Plant_Emergence$Rep <- as.factor(Plant_Emergence$Rep)


str(Plant_Emergence)
```


2. 5 pts. Fit a linear model to predict Emergence using Treatment and
DaysAfterPlanting along with the interaction. Provide the summary of the linear
model and ANOVA results.


```{r}

# We can make the model like this
lm1 <- lm(Emergence~Treatment + DaysAfterPlanting + Treatment:DaysAfterPlanting, data = Plant_Emergence)

# Or like this, these are the same thing:
lm1 <- lm(Emergence~Treatment * DaysAfterPlanting, data = Plant_Emergence)

summary(lm1)
anova(lm1)

```


3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

  Based on the results, no, it does not look like we need to fit the interaction term, as the interaction between `Treatment` and `DaysAfterPlanting` was not significant for the anova.  



```{r}

# Or like this, these are the same thing:
lm1 <- lm(Emergence~Treatment * DaysAfterPlanting, data = Plant_Emergence)

summary(lm1)
anova(lm1)

```







