---
title: "coding_challenge_6"
author: "AS-SL"
date: "2025-03-27"
output: html_document
---
1. 2 pts. Regarding reproducibility, what is the main point of writing your own functions
and iterations?


When we are dealing with big sets of data were we have to perform repetitive actions or mathemathical operations, it is better to have a function or iteration to go trough all the data instead of copying and pasting the information. It also makes it easier for someone else to understand what we are doing. Writing functions and using iterations helps us understand what we didi as well so that we can re-run our data or use the same analysis for a different set of data.

2. 2 pts. In your own words, describe how to write a function and a for loop in R and how
they work. Give me specifics like syntax, where to write code, and how the results are
returned.

To write a function first we give our function a name that could be related to what it does , then we add the variable in a parentheses. We call for the function definition of Rstudio and write down what we want our function to do using code inside {}.Finally we add the return() output of the function 
```{r}
name<-function(variable){code
  return()}
```

To write for loop in R, the common synthax 
  1.- for()
 2. i common name or a varible
 3. in (sequence we want to iterate)
 4. {} to tell it what to do with that input
 5. function
 6.print or show the output in some way

```{r}
for (i in sequence we want to iterate){
 funtion 
  output 
}
```

Load packages

```{r}
library(readr)
library(tidyverse)

```



This dataset contains the population and coordinates (latitude and longitude) of the 40 most populous cities in the US, along with Auburn, AL. Your task is to create a function that calculates the distance between Auburn and each other city using the Haversine formula. To do this, you'll write a for loop that goes through each city in the dataset and computes the distance from Auburn. Detailed steps are provided below.

3. 2 pts. Read in the Cities.csv file from Canvas using a relative file path.
```{r}
Cities_US<- read.csv("Cities.csv")
View(Cities_US)
```

4. 6 pts. Write a function to calculate the distance between two pairs of coordinates based on the Haversine formula (see below). The input into the function should be lat1, lon1, lat2, and lon2. The function should return the object distance_km. All the code below
needs to go into the function.

Create function 

```{r}
mad_function<- function(lat1,lon1,lat2,lon2){# convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180
# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a))
# Earth's radius in kilometers
earth_radius <- 6378137
# Calculate the distance
distance_km <- ((earth_radius * c)/1000)
return(distance_km)}
```

test the function

```{r}
mad_function(32.6087,-85.4903,40.6943,-73.9249)
```


5. 5 pts. Using your function, compute the distance between Auburn, AL and New York
City
a. Subset/filter the Cities.csv data to include only the latitude and longitude values
you need and input as input to your function.


sub-setting the csv file to include only the latitude and longitude values needed
```{r}

NY<-Cities_US%>%
  select(city,lat,long)%>%
  filter(city=="Auburn" | city == "New York")
NY
```
b. The output of your function should be 1367.854 km
```{r}
mad_function(NY[1,2],NY[1,3],NY[2,2],NY[2,3])
```


6. 6 pts. Now, use your function within a for loop to calculate the distance between all other
cities in the data. The output of the first 9 iterations is shown below.
## [1] 1367.854
## [1] 3051.838
## [1] 1045.521
## [1] 916.4138
## [1] 993.0298
## [1] 1056.022
## [1] 1239.973
## [1] 162.5121
## [1] 1036.99
Bonus point if you can have the output of each iteration append a new row to a
dataframe, generating a new column of data. In other words, the loop should create a
dataframe with three columns called city1, city2, and distance_km, as shown below. The
first six rows of the dataframe are shown below.
## City1 City2 Distance_km
## 1 New York Auburn 1367.8540
## 2 Los Angeles Auburn 3051.8382
## 3 Chicago Auburn 1045.5213
## 4 Miami Auburn 916.4138
## 5 Houston Auburn 993.0298
## 6 Dallas Auburn 1056.0217

```{r}

city_names<-unique(Cities_US$city)

filtered_city<-Cities_US%>%
  select(city,lat,long)%>%

for (i in seq_along(city_names)){
  
}
  
```

7. 2 pts. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6.
Provide me a link to your github written as a clickable link in your .pdf or .docx

 



 
